Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-27 TACAS2018/paper.tex
Moved property updates to corresponding sections

The CEGIS-OR loop for this approach is described next.
We make use of two abstractions corresponding to two
specifications: $\phi_{synth}$ used by the synthesiser and
$\phi_{static}$ which is given to the synthesis refinement loop. 
Both specifications will be updated during execution and new abstractions will be constructed for each.
Let us define the properties 
\begin{align*}
\phi_{FWL}^C&:= \mat{K}=\mat{K}_{\langle C \rangle} \wedge \mat{L}=\mat{L}_{\langle C \rangle}\\
\phi_{sample}^{d}&:= \mat{A}_d, \mat{B}_d, \mat{C}_d
\end{align*}
We start our loop with 
\begin{align*}
\phi_{synth}&:=\phi_{sample}^{d} \wedge \phi_{stability} \wedge \phi_{input} \wedge \phi_{init}\\
\phi_{static}&:=\phi_{synth} \wedge \phi_{FWL}^C.
\end{align*}
When a counterexample is found, we add it to $\phi_{synth} = \phi_{synth} \wedge \phi_{safety}^k$, 
where k is the iteration found for the counterexample.  The ${\sc Abstract}$ phase will then construct a abstraction (in the form of a program) that meets this property (unsoundly since the program uses FWL).
It is also possible, when the modeller is unable to increase $k$, to change $\phi_{synth} = \phi_{synth} \wedge \phi_{sample}^{d'}$ or $\phi_{synth} = \phi_{synth} \wedge \phi_{FWL}^C$.  What we point out here is that these specifications may grow up to the full specification, thus making for increasingly complex astractions.
Finally, we also note that the abstract phase uses a given precision ($\phi_{FWL}^P$) which we will increase when no candidates can be generated for a given $\phi_{synth}$.
In the modeller, $\phi_{static} = \phi_{static} \wedge \phi_{safety}^{k'} \wedge \phi{objective}$ where $k' \geq k$, and $\phi{objective}$ defines a set of rules for the current optimization stage (see {\sc Opimise}).  We only note that at all times $\phi_{static} \models \phi_{synth}$.
Let us see how the loop updates these specifications.

\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:observable}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

@@

If we cannot optimize the candidate such that it passes the {\sc
  verify synthesis} check, we signal the {\sc Abstract} phase to
refine $\phi_{synth}$ and request {\sc Synthesise} to produce a new
candidate.  Otherwise the refined candidate solution is given back to
{\sc Verify Synthesis}.
%
%\item
\paragraph{5. {\sc Discrete Time}} verifies $\phi_{safety}^{\infty}$ (unbounded time)
using discrete time dynamics.  If the verification fails, we report the
counterexample to {\sc Refine Abstraction} including details of initial states
and iterations that failed the specification.  
%\item
\paragraph{6. {\sc Refine Abstraction}} changes $\phi_{safety}$ to encompass any arising
counterexample.  This can be done by increasing the time horizon $k$ for the
safety specification of the violating initial condition and input.
It also modifies $\phi_{static}$ to account for the counterexample.
If the magnitude of the violation is small, {\sc Optimise} may find a new
candidate, otherwise it will feed through to {\sc Abstract}.
%
%\item
\paragraph{7. {\sc Continuous Time}} verifies that the
transitions between sample times meet the specification $\phi_{safety}^{\infty}$.  If
a counterexample is found, information about the maximum distance of the
violation to the specification boundary is fed to {\sc Sample}. 
%\item
\paragraph{8. {\sc Sample}} selects a smaller suitable sample time and modify the abstractions to meet the discretization. 
It will then feed back to the abstraction refinement loop to look for a suitable action for the updated model.
%

%\item
If we reach the {\sc abstract} phase, it means that the candidate
  solution is not valid, in which case we must refine $\phi_{synth}$. If $\phi_{synth}$ cannot be further refined, 
  then we cannot synthesise a new candidate and return a synthesis failure.
%\end{enumerate}

@@

The benchmark suite employed to evaluate the proposed technique is 
composed by seven SISO \note{defined SISO} models extracted from the control literature~\cite{acrobot,cstr,CHEN1979389,KOKOTOVIC198023,gajic2008optimal,Franklin15,maglev,converters}.
In order to obtain statistically relevant results, we have taken these models
and modified each possible combination of their parameters by a 10\% tolerance,
thus creating seven hundred benchmarks from this base.
In the verification module, we have used the tool Axelerator~\cite{} .

>>>>

The CEGIS-OR loop for this approach is described next.
We make use of two abstractions corresponding to two
specifications: $\phi_{synth}$ used by the synthesiser and
$\phi_{static}$ which is given to the synthesis refinement loop. 
Both specifications will be updated during execution and new abstractions will be constructed for each.
Let us define the properties 
\begin{align*}
\phi_{\langle C \rangle}&:= \mat{K}=\mat{K}_{\langle C \rangle} \wedge \mat{L}=\mat{L}_{\langle C \rangle}\\
\phi_{dynamics}^{T_d}&:= \mat{A}_d, \mat{B}_d, \mat{C}_d
\end{align*}
The first one defines the word length of the digital controller ($\mathbb{R}_{\langle C \rangle}$) and its effect on the system dynamics.
The second one defines the sampling time ($T_d$) and its effect on the system dynamics (as per Section \ref{sec:continuous_time_accel}).
We start our loop with 
\begin{align*}
\phi_{synth}&:=\phi_{dynamics}^{T_d} \wedge \phi_{stability} \wedge \phi_{input} \wedge \phi_{init}\\
\phi_{static}&:=\phi_{synth} \wedge \phi_{\langle C \rangle}.
\end{align*}
These specifications will vary as the loop progresses.  We only note that at all times $\phi_{static} \models \phi_{synth}$.
Let us see how the loop updates these specifications.

\paragraph{1. {\sc abstract}} creates an abstraction from the initial specification as follows:
\begin{itemize}
\item Select a sample time ($T_d$) that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep2}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. 
%The transformation matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:cof_fwl_stability}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
When a counterexample is found, we add it to $\phi_{synth} = \phi_{synth} \wedge \phi_{safety}^k$, 
where k is the iteration found for the counterexample.  The ${\sc Abstract}$
phase will then construct a abstraction (in the form of a program) that meets
this property (unsoundly since the program uses FWL).

%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision ($\phi_{\langle P \rangle}$), or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

@@

We denote these sets of rules as $\phi_{objective}$ (which varies depending on which optimization stage we're in).
At each optimizing iteration, the modeller updates $\phi_{static} = \phi_{static} \wedge \phi_{objective}$% \wedge \phi_{safety}^{k'} $ where $k' \geq k$.

If we cannot optimize the candidate, we signal the {\sc Abstract} phase to
refine $\phi_{synth}$ and request {\sc Synthesise} to produce a new
candidate.  Otherwise the optimised candidate solution is given back to
{\sc Verify Synthesis}.

%\item
\paragraph{5. {\sc Discrete Time}} verifies $\phi_{safety}^{\infty}$ (unbounded time)
using discrete time dynamics.  If the verification fails, we report the
counterexample to {\sc Refine Abstraction} including details of initial states
and iterations that failed the specification.  
%\item
\paragraph{6. {\sc Refine Abstraction}} changes $\phi_{static} = \phi_{static} \wedge \phi_{safety}^k$
to encompass any arising counterexample.  This is done by increasing the time horizon $k$ for the
safety specification of the violating initial condition and input.
If the magnitude of the violation is small, {\sc Optimise} may find a new
candidate, otherwise it will feed through to {\sc Abstract}.
%
%\item
\paragraph{7. {\sc Continuous Time}} verifies that the
transitions between sample times meet the specification $\phi_{safety}^{\infty}$.  If
a counterexample is found, information about the maximum distance of the
violation to the specification boundary is fed to {\sc Sample}. 
%\item
\paragraph{8. {\sc Sample}} selects a smaller suitable sample time and modify the abstractions to meet the discretization 
($\phi_{synth} = \phi_{synth} \wedge \phi_{sample}^{T_d'}$ and $\phi_{static} = \phi_{static} \wedge \phi_{sample}^{T_d'}$).
It will then feed back to the abstraction refinement loop to look for a suitable action for the updated model.
%\end{enumerate}

@@

The benchmark suite employed to evaluate the proposed technique is 
composed by seven models extracted from the control literature~\cite{acrobot,cstr,CHEN1979389,KOKOTOVIC198023,gajic2008optimal,Franklin15,maglev,converters}.
In order to obtain statistically relevant results, we have taken these models
and modified each possible combination of their parameters by a 10\% tolerance,
thus creating seven hundred benchmarks from this base.
In the verification module, we have used the tool Axelerator~\cite{} .
